---
title: '[STAT 4540] HW-4'
author: "Michael Ghattas"
date: "3/16/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

### (a)

```{r}
library(lubridate)
library(gridExtra)
library(ggplot2)
library(dplyr)

df <- load("/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4540/HW/4/HW4_problem1_data.RData");
dat <- data.frame(x = x)
p = 3

phi = ar.yw(dat, aic = FALSE, order.max = p)

datACF = acf(dat, p, type = "covariance")
gam <- datACF[1:p]
g1 = as.numeric(datACF[1])[1]
g2 = as.numeric(datACF[2])[1]
g3 = as.numeric(datACF[3])[1]
gamma = c(g1, g2, g3)
g0 = as.numeric(datACF[0])[1]
g1 = as.numeric(datACF[1])[1]
g2 = as.numeric(datACF[2])[1]
g = c(g0, g1, g2)
Gamma = toeplitz(g)

phi_hat = solve(Gamma, gamma)

phi
phi_hat
```

### (b)

```{r}
sigma2_hat = g0 - (gamma %*% phi_hat)
sigma2_hat
```

### (c)

```{r}
AR <- arima.sim(list(ar = c(-1, -0.1)), 100)

mod1 = ar.yw(AR, order.max = 1, aic = FALSE)
mod2 = ar.yw(AR, order.max = 2, aic = FALSE)
mod99 = ar.yw(AR, order.max = 99, aic = FALSE)

predict(mod1)$pred[1]
predict(mod2)$pred[1]
predict(mod99)$pred[1]
```
```{r}
p1 = rep(0, 1000)
p2 = rep(0, 1000)
p99 = rep(0, 1000)

for (i in 1:1000)
{
  AR <- arima.sim(list(ar = c(-1, -0.1)), 101); AR
  
  mod1 = ar.yw(AR[1:100], order.max = 1, aic = FALSE)
  mod2 = ar.yw(AR[1:100], order.max = 2, aic = FALSE)
  mod99 = ar.yw(AR[1:100], order.max = 99, aic = FALSE)
  
  pred1 = predict(mod1)$pred[1]
  pred2 = predict(mod2)$pred[1]
  pred99 = predict(mod99)$pred[1]
  
  p1[i] = pred1 - AR[101]
  p2[i] = pred2 - AR[101]
  p99[i] = pred99 - AR[101]
}

mean(p1); mean(p2); mean(p99)
sd(p1); sd(p2); sd(p99)
```
##### With enough iterations we can start to see a trend of increasing values for the average prediction error and standard deviation.

# Problem 2

### (a)

$X_{t} = Z_{t} + \theta \cdot Z_{t - 1}$ \
$Z_{t} \sim WN(0, \sigma^2)$ \

$a = \Gamma^{-1} \cdot \gamma$ \
$\hat{x_{3}} = a_{1} \cdot x_{1} + a_{2} \cdot X_{2} + a_{3} \cdot X_{4} + a_{4} \cdot X_{5}$ \
$W = (X_{1}, X_{2}, X_{4}, X_{5})$ \

$\gamma = [cov(X_{3}, X_{1}), cov(X_{3}, X_{2}), cov(X_{3}, X_{4}), cov(X_{3}, X_{5})] = [\gamma(2), \gamma(1), \gamma(1), \gamma(2)] = [0, \theta \cdot \sigma^2, \theta \cdot \sigma^2, 0]$ \

$\Gamma =$ \
$|var(x_{1}), cov(x_{1}, x_{2}), cov(x_{1}, x_{4}), cov(x_{1}, x_{5})|$ \
$|cov(x_{2}, x_{1}), var(x_{2})), cov(x_{2}, x_{4}), cov(x_{2}, x_{5})|$ \
$|cov(x_{4}, x_{1}), cov(x_{4}, x_{2}), var(x_{4})), cov(x_{4}, x_{5})|$ \
$|cov(x_{5}, x_{1}), cov(x_{5}, x_{2}), cov(x_{5}, x_{4}), var(x_{5})|$ \
$=$ \
$|\gamma(0), \gamma(1), \gamma(3), \gamma(4)|$ \
$|\gamma(1), \gamma(0), \gamma(2), \gamma(3)|$ \
$|\gamma(3), \gamma(2), \gamma(0), \gamma(1)|$ \
$|\gamma(4), \gamma(3), \gamma(1), \gamma(0)|$ \
$=$ \
$|\sigma^{2} \cdot (1  + \theta^{2}), \theta \cdot \sigma^2, 0, 0|$ \
$|\theta \cdot \sigma^2, \sigma^{2} \cdot (1  + \theta^{2}), 0, 0|$ \
$|0, 0, \sigma^{2} \cdot (1  + \theta^{2}), \theta \cdot \sigma^2|$ \
$|0, 0, \theta \cdot \sigma^2, \sigma^{2} \cdot (1  + \theta^{2})|$ \

$\Gamma^{-1} =$ \
$|\frac{1}{\sigma^{2} \cdot (1  + \theta^{2})}, \frac{1}{\theta \cdot \sigma^2}, 0, 0|$ \
$|\frac{1}{\theta \cdot \sigma^2}, \frac{1}{\sigma^{2} \cdot (1  + \theta^{2})}, 0, 0|$ \
$|0, 0, \frac{1}{\sigma^{2} \cdot (1  + \theta^{2})}, \frac{1}{\theta \cdot \sigma^2}|$ \
$|0, 0, \frac{1}{\theta \cdot \sigma^2}, \frac{1}{\sigma^{2} \cdot (1  + \theta^{2})}|$ \

$\hat{X_{3}} = P(X_{3} | W) = a^{T} \cdot W = \Gamma^{-1} \cdot \gamma \cdot [X_{1}, X_{2}, X_{4}, X_{5}]^{T}$ \

### (b)

```{r}
sig2 = (1/3)
theta = sqrt(2)

gamma0 = sig2 * (1 + theta^2)
gamma1 = sig2 * theta
gam = c(0, gamma1, gamma1, 0)

gc1 = c(gamma0, gamma1, 0, 0)
gc2 = c(gamma1, gamma0, 0, 0)
gc3 = c(0, 0, gamma0, gamma1)
gc4 = c(0, 0, gamma1, gamma0)
Gam = as.matrix(cbind(gc1, gc2, gc3, gc4))

a = as.matrix(solve(Gam, gam))

MSE = gamma0 - (t(a) %*% gam); MSE

```

# Problem 3

### (a)

If $X \sim N(0, \Sigma) \to X \sim N(\vec{0}, \sigma^{2}I)$, this means $x_{i}$ and $x_{j}$ are uncorrelated and independent for MVN. Therefore $Y = \Sigma^{-\frac{1}{2}}X$ is $Y \sim MVN(\mu, \Sigma)$ and iid.

### (b)

$f(X) = f(X_{1}, X_{2}, X_{3}, X_{4}) = f(\vec{X})$ \
Joint distribution for $X ~ \sim N(\mu, \Sigma) = f(\mu, \theta, \sigma^{2})$ \
$f(\mu, \theta, \sigma^{2}) = \prod_{i = 1}^{4} \frac{1}{2\pi \cdot |\Sigma^{-\frac{1}{2}}|} \exp{-\frac{1}{2} \cdot (\vec{x} - \vec{\mu}) \cdot \Sigma^{-\frac{1}{2}} \cdot (\vec{x} - \vec{\mu})}$ \ 

Where:

$\vec{x} = [x_{1}, x_{2}, x_{3}, x_{4}]$ \
$\vec{\mu} = [0, 0, 0, 0]$ \
$\Sigma^{-\frac{1}{2}} =$ \
$|\sqrt{var(x_{1})}, \sqrt{cov(x_{1}, x_{2})}, \sqrt{cov(x_{1}, x_{3})}, \sqrt{cov(x_{1}, x_{4})}|$ \
$|\sqrt{cov(x_{2}, x_{1})}, \sqrt{var(x_{2})}, \sqrt{cov(x_{2}, x_{3})}, \sqrt{cov(x_{2}, x_{4})}|$ \
$|\sqrt{cov(x_{3}, x_{1})}, \sqrt{cov(x_{3}, x_{2})}, \sqrt{var(x_{3}))}, \sqrt{cov(x_{3}, x_{4})}|$ \
$|\sqrt{cov(x_{4}, x_{1})}, \sqrt{cov(x_{4}, x_{2})}, \sqrt{cov(x_{4}, x_{3})}, \sqrt{var(x_{4})}|$ \
$=$ \
$|\sqrt{\gamma(0)}, \sqrt{\gamma(1)}, \sqrt{\gamma(2)}, \sqrt{\gamma(3)}|$ \
$|\sqrt{\gamma(1)}, \sqrt{\gamma(0)}, \sqrt{\gamma(1)}, \sqrt{\gamma(2)}|$ \
$|\sqrt{\gamma(2)}, \sqrt{\gamma(1)}, \sqrt{\gamma(0)}, \sqrt{\gamma(1)}|$ \
$|\sqrt{\gamma(3)}, \sqrt{\gamma(2)}, \sqrt{\gamma(1)}, \sqrt{\gamma(0)}|$ \
$=$ \
$|\sigma\sqrt{1 + \theta_{1}^{2} + \theta_{2}^{2}}, \sigma\sqrt{\theta_{1} \cdot (1 + \theta_{2})}, \sigma\sqrt{\theta_{2}}, 0|$ \
$|\sigma\sqrt{\theta_{1} \cdot (1 + \theta_{2})}, \sigma\sqrt{1 + \theta_{1}^{2} + \theta_{2}^{2}}, \sigma\sqrt{\theta_{1} \cdot (1 + \theta_{2})}, \sigma\sqrt{\theta_{2}}|$ \
$|\sigma\sqrt{\theta_{2}}, \sigma\sqrt{\theta_{1} \cdot (1 + \theta_{2})}, \sigma\sqrt{1 + \theta_{1}^{2} + \theta_{2}^{2}}, \sigma\sqrt{\theta_{1} \cdot (1 + \theta_{2})}|$ \
$|0, \sigma\sqrt{\theta_{2}}, \sigma\sqrt{\theta_{1} \cdot (1 + \theta_{2})}, \sigma\sqrt{1 + \theta_{1}^{2} + \theta_{2}^{2}}|$ \
