---
title: '[STAT 4540] HW-1 / Michael Ghattas'
author: "Michael Ghattas"
date: "1/26/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

### a)

$X_t = \{X_1, X_2, X_3\}$ \

$\{X_1, X_2, X_3\} \to independent$ \

$X_1 \sim Poisson(1, 2)$ \

$X_2 \sim Expo(1, 0.5)$ \

$X_3 \sim \Gamma(1, 0.1)$ \

### b)

$X_t \sim^{iid} Normal(0, 1)$ $\forall t$ \

$t = \{0, \infty\}$ \


# Problem 2

### a)

```{r}
set.seed(1)



U_1 <- rnorm(1,0,1)
U_2 <- rnorm(1,0,1)
U_3 <- rnorm(1,0,1)
U <- c(U_1, U_2, U_3)


V_1 <- rnorm(1,0,1)
V_2 <- rnorm(1,0,1)
V_3 <- rnorm(1,0,1)
V <- c(V_1, V_2, V_3)


t <- seq(0, (2 * pi), by = 0.01)


y_1 <- ((U_1 * sin(t)) + (V_1 * cos(t)))
y_2 <- ((U_2 * sin(t)) + (V_2 * cos(t)))
y_3 <- ((U_3 * sin(t)) + (V_3 * cos(t)))
Y <- c(y_1, y_2, y_3)



plot(t, y_1, ylim = c(-(2 * pi), (2 * pi)), type = 'l')
points(t , y_2, col = "red", pch = ".")
points(t, y_3, col = "blue", pch = ".")
```

### b)

##### Using the given information:

$E[X_t] = E[U] * \sin{(t)} + E[V] * \cos{(t)} = 0$ \

$E[X_s] = E[U] * \sin{(s)} + E[V] * \cos{(s)} = 0$ \

We will test if $E[X_t, X_s] = E[X_{t + h} \cdot X_{s + h}]$ $\forall (h)$ & $\forall (t, s)$: \
$E[X_t \cdot X_s] = (E[U] * \sin{(t)} + E[V] * \cos{(t)}) * (E[U] * \sin{(s)} + E[V] * \cos{(s)})$ \
                 $= (U * \sin{(t)} * U \sin{(s)}) + (U * \sin{(t)} * V * \cos{(s)}) + (V * \cos{(t)} * U * \sin{(s)}) + (V * \cos{(t)} * V * \cos{(s)})$ \
                 $= U^2 \cdot \sin{(t)}\sin{(s)}) + UV \cdot \sin{(t)}\cos{(s)} + UV \cdot \cos{(t)}\sin{(s)}) + V^2 \cdot \cos{(t)}\cos{(s)})$ \
                 $= 0.5U^2\sin{0.5(t + s)}\sin{0.5(t - s)} + 0.5V^2\cos{0.5(t + s)}\cos{0.5(t - s)}) + UV\sin{(t + s)}$ \

Since the ACVF will not be only dependent on $(t - s)$, $\{X_t\}$ is NOT weakly stationary! \



# Problem 3

### a

Since $Y_0 = 0$ is a provided condition, this is not a standard condition of a n AR(1) model. This means we set a limitation on the starting point of the series. \


### b

$y_t = \phi \cdot y_{t - 1} + z_t$ \

$E[y_t] = E[\phi \cdot y_{t - 1} + z_t] = \phi \cdot E[y_{t - 1}] + E[z_t] = \phi \cdot \mu + 0 = \phi \cdot \mu$ \

$\mu = \phi\mu$ \

$\mu - \phi\mu = 0$ \

$\mu \cdot (1 - \phi) = 0$ \

$\mu = 0$ \

$E[y_t] = \mu = 0$ \

### c

$Var[y_t] = E[y_{t - 1}^{2}] - E[y_t]^2 = E[y_{t - 1}^{2}] - 0^2 = E[(\phi \cdot y_{t - 1} + z_t)^2] - 0$ \
         $= E[(\phi y_{t - 1})^2 + 2 \cdot \phi y_{t - 1} \cdot z_1 + z^2] = E[(\phi)^2] \cdot E[(y_{t - 1})^2] + 2\phi(0)(0) + 0^2]$ \
         $= \phi^2E[(y_{t - 1})^2] + 0 + 0 = \phi^2E[(0)^2] = \phi^2(0) = 0$ \

### d

##### Using the information we have so far:

1. $E[y_t] = \mu = 0$ \

2. $Var[y_t] = 0 \to \sigma = 0$ \

3. $E[y_t, y_s] = E[y_{t + h} \cdot y_{s + h}] = E[y_0y_{s - t}$, where $(h = -t)$ \

   $E[y_t \cdot y_s] = 0$, given $(y_0 = 0)$ \
   
Thus, $\{y_t\}$ is weakly stationary! \


# END.